Estimating Correspondences of Deformable Objects ``In-the-wild"

### 2016-04-09 contents/implementation.tex
cam ready final draft

This section presents the proposed method for establishing dense correspondences among training shapes
by only using curve annotations.
%
%We propose a novel framework for building deformable models that does not require any consistent set of annotated landmarks and is based on a dense shape representation. Our method only requires a set of point or curve line annotations that does not need to be consistent over different training samples.
%
%It combines the techniques of Non-rigid Iterative Closest Point (NICP) \cite{Amber2007}, multi-channel Support Vector Shape (SVS) \cite{Nguyen2013} representation and multi-image subspace flow \cite{Garg:2013hu} in an effective framework that has significant descriptive power.
%
It takes as input a set of training images of a particular object class, along with the corresponding curve annotations. The steps of our pipeline, which are also depicted in Figure~\ref{fig:archi}, are the following:

@@

\noindent\textbf{Step 2:} Construct a correspondence basis for the training SVS images. This is acquired by applying Non-rigid ICP on the densely sampled annotated curves, followed by Principal Component Analysis (PCA).

@@

In order to fully capture the variability among most deformable objects' shapes annotations, we use a representation based on SVS \cite{Nguyen2013}. An SVS is a decision function trained on shapes using Support Vector Machines (SVMs) with Radial Basis Function (RBF) kernels. In this way, a shape is represented as a classifier function, which has several advantages: (a)~the representation is completely generic, e.g. it can be applied to sparse landmark points, curves lines or a combination of the two, and~(b) it fuses inconsistent landmarks into consistent and directly comparable decision functions. Furthermore, this representations is also robust against noise, missing data and outliers \cite{Nguyen2013}.

@@

We extend the SVS representation to support also the case where multiple curves with different labels are annotated. This is useful when annotating different structures of the same object, such as the left and right eyebrows on faces. Even though not absolutely necessary in our framework, it can provide further guidance on the estimation of dense shape correspondences for various object classes. In more detail, we create a multi-channel SVS image $\bm{d}(\bm{x})=[d_1(\bm{x}) \cdots d_i(\bm{x}) \cdots d_{N_c}(\bm{x})]$, where $d_i(\bm{x})$ is the SVS image that corresponds to the curve annotation of the $i$-th structure and $N_c$ is the total number of structures (a single curve annotation is the special case where $N_c=1$). Note that we do not necessarily require that all structures are annotated in all the object images: in the case that a structure is not annotated, the corresponding channel of the SVS image simply takes a zero value for all pixels. The shape flow estimation can deal with such missing information thanks to the spatial regularisation and the low-rank constraint that it adopts, c.f.~Step~\ref{sec:step3}.

@@

Instead of the \emph{motion} basis used in multiframe optical flow formulation of \cite{Garg:2013hu}, we build a \emph{correspondence} basis that introduces constraints on how points of different shapes are matched to each other. Every pixel of the reference SVS image is matched to its corresponding position at every training SVS image and in this way defines a \emph{correspondence vector}. This vector consists of the 2D locations of the specific point in all SVS images. To form this vector, the training images are arranged in an arbitrary order. Similarly to the order of training samples when PCA is applied, this order does not affect the result of our method and any re-ordering would produce exactly the same results.

@@

Using the constructed correspondence basis, the shape flow can be approximated as:
\vspace{-5pt}
\begin{equation}\label{eq:LinTrajModel}
    \bm{u}_n(\bx) \approx
    \textstyle \, \sum_{i=1}^R \, \bm{q}_i(n)\bm{v}_i(\bx) \,\, ,
\vspace{-5pt}
\end{equation}
where $\bm{v}_i(\bx)$ is the weight that needs to be applied on the $i$-th correspondence basis element, in order to get the correspondence vector for the point location $\bm{x}$. In other words, the shape flow for every point $\bm{x}$ is described as a linear combination of basis elements that is controlled by the coefficients $\bm{v}_i(\bm{x})$.
The values of the $i$-th coefficient for all the points $\bm{v}_i(\bm{x})$ can be interpreted as an image defined on $\Omega$. Using the correspondence basis, the determination of the shape flow boils down to the determination of the set of coefficients $\bm{v}_i(\bm{x})$. The above representation of shape flow, constrains the correspondence vectors to lie on a subspace and, therefore, acts as a low-rank prior that enforces coherency of the shape registration result over the whole training dataset of shapes.

To effectively build the correspondence basis, we first transform the original annotations to sparse point clouds. Then, we apply Nonrigid Iterative Closest Point (NICP) \cite{Amber2007} between the point cloud of annotations in the reference shape and the one of every shape of the training set. NICP iteratively deforms the cloud of points of every shape to match the points of the reference shape. This yields an initial estimation of the correspondence vectors on the sparse locations of annotated landmarks on the reference shape. Finally, the correspondence basis is found by applying PCA on these correspondence vectors and keeping only the first $R$ principal components.

@@

\begin{figure}[b!]
    \centering
    \includegraphics[width=0.9\columnwidth]{resources/models}
    \caption{Dense shape models build for faces and ears. Dense shapes are presented as grid for better visualisation.}
    \label{fig:dense_models}
\end{figure}

@@

Let $\bm{d}(\bx;n)$, $\bm{d}(\bx;0):\Omega\rightarrow \R^{N_c}$ be the $n$-th training SVS image and the reference SVS image respectively. Following \cite{Garg:2013hu}, we propose to estimate the shape flow over all training images by minimizing the following energy:
\vspace{-5pt}
\begin{align}
E_{sf} & =\alpha
\int_{\Omega}\sum_{n=1}^{N_t} \|\bm{d}(\bx+\bm{u}_n(x);n)-\bm{d}(x;0)\| \ud \bx \label{eq:costfunc}\\
    &+ \beta \int_{\Omega}\sum_{n=1}^{N_t}\|\bm{u}_n(\bx)-\sum_{i=1}^R\bm{q}_i(n)\bm{v}_i(\bx)\|^2 \ud \bx \label{eq:lowrank}\\
    &+
\int_\Omega  \sum_{i=1}^R \,\, \left \|    \nabla \bm{v}_i(\bx)    \right \|  \,\ud \bx \label{eq:TVterm}
\vspace{-5pt}
\end{align}
This energy consists of two sets of unknown shape flows that are relatively close to each other: (i) $\bm{u}_n(x)$ which tries to explain the data from the input SVS images, and (ii) the shape flow determined by the correspondence basis coefficients $\bm{v}_i(\bx)$ that are spatially regularised and enforce a low-rank prior.
%We minimize this energy jointly with respect to $\bm{u}_n(\bx)$ and $\bm{v}_i(\bx)$.
%The positive parameters $\alpha$ and $\beta$ weigh the balance between the terms of the energy.

The \textbf{first term} of the above energy \eqref{eq:costfunc} is a data attachment term
that uses the robust $\Lone$-norm.  It is based on the assumption that the values of the reference SVS image $\bm{d}_0(\bx)$ at every pixel $\bx$ are preserved at its corresponding locations on all training SVS images $\bm{d}_n(\bx)$. The use of an $\Lone$-norm improves the robustness of the method since it allows deviations from this assumption, which might occur in practice.
The \textbf{second term} of the energy \eqref{eq:lowrank} penalizes the difference between the two sets  of shape flows and acts as a coupling term between them.
The \textbf{third term} of the energy \eqref{eq:TVterm} corresponds to the spatial Total Variation regularization \cite{rudin92} of
the correspondence basis coefficients $\bm{v}_i(\bx)$.
This term penalizes spatial oscillations of each coefficient caused by distortions of the SVS images but not strong discontinuities that are desirable in the borders of different object regions. In addition, this term allows to fill in information into regions where the shape information in the SVS images is missing, due to e.g.~regions with no annotations.

@@

%In every coarse-to-fine and warping iteration, we use an initialization that comes from the previous iteration and for the very beginning, we initialize with a Thin Plate Splines (TPS) \cite{Bookstein1989} interpolation of the initial correspondence vectors described in Sec.~\ref{sec:step2}. We approximate the data term \eqref{eq:costfunc} by linearizing the SVS images around the initialization. After that, the energy becomes convex and we optimize it using alternating optimization w.r.t.~$\bm{v}_i(\bx)$ and $\bm{u}_n(\bx)$. The minimization w.r.t.~
%$\bm{v}_i(\bx)$ is decoupled for every coefficient $i$ and corresponds to Rudin-Osher-Fatemi Total Variation denoising \cite{rudin92}, which we solve efficiently by applying the first order primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}. The minimization w.r.t.~
%$\bm{u}_n(x)$ is decoupled for every pixel $\bx$ and every shape index $i$. This minimization is also implemented by applying the efficient primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}.

@@

Given a test image $\mathbf{I}$, the fitting process involves the minimisation of the following cost function:
\begin{equation}
    \arg\min_{\bm{p}, \bm{c}}\|\bm{I}(\mathcal{W}(\bm{s}(\bm{p}))) - \bm{t}(\bm{c})\|_2^2
	\label{eq:aam_cost}
\end{equation}
This optimisation problem is typically solved using the inverse-compositional Gauss-Newton algorithm, for which different variations have been proposed \cite{Matthews2004, Papandreou2008, Amberg2009, Tzimiropoulos2013, Alabort2014}. Note that the existence of the sampling function $\mathcal{W}()$ instead of a non-linear warping function has the advantage that all existing efficient gradient descent algorithms become exact.

\paragraph{Outline Patch-Based Active Appearance Model} The object classes for which the interior appearance does not have specific structure are modeled using patch-based AAMs \cite{Tzimiropoulos2014} trained on a set of sparse landmarks. Especially for human body parts (arms, legs), we strongly believe that the points located to the outline of the object are more suitable compared to the internal ones that correspond to the skeleton joints, which are commonly used by current literature \cite{buehler2011upper,charles2013domain,pfister2015flowing,yang2013articulated}.

The main differences between the patch-based and dense AAMs are that (a)~the densified shape instances are sub-sampled to include only the outline points, and (b)~the texture representation involves the sampling a neighbourhood around each point instead of a single pixel. Specifically, in order to build the outline sparse shape model, we simply select the outline points on the SVS reference frame. Then, by taking advantage of the dense correspondences obtained by Step \ref{sec:step3}, the shape model is trained in a similar way as in the dense case. Moreover, similar to the dense case, the texture model is built by sampling the image values from the sparse shape locations, i.e. $\mathcal{W}(\bm{s}(\bm{p}))$. However, contrary to dense AAMs, we sample a patch that is centred around each landmark point. These patches are then vectorised and concatenated in a single texture vector. Note that the optimisation process remains exactly the same.

>>>>

This section presents the proposed method for establishing dense correspondences among training shapes
by only using curve annotations.
%
%We propose a novel framework for building deformable models that does not require any consistent set of annotated landmarks and is based on a dense shape representation. Our method only requires a set of point or curve line annotations that does not need to be consistent over different training samples.
%
%It combines the techniques of Non-rigid Iterative Closest Point (NICP) \cite{Amber2007}, multichannel Support Vector Shape (SVS) \cite{Nguyen2013} representation and multiimage subspace flow \cite{Garg:2013hu} in an effective framework that has significant descriptive power.
%
It takes as input a set of training images of a particular object class, along with the corresponding curve annotations. The steps of our pipeline, which are also depicted in Figure~\ref{fig:archi}, are the following:

@@

\noindent\textbf{Step 2:} Construct a correspondence basis for the training SVS images. This is acquired by applying the Non-rigid ICP (NICP) algorithm of \cite{Amber2007} on the densely sampled annotated curves, followed by Principal Component Analysis (PCA).

@@

In order to fully capture the variability among most deformable objects' shapes annotations, we use a representation based on SVS \cite{Nguyen2013}. An SVS is a decision function trained on shapes using Support Vector Machines (SVMs) with Radial Basis Function (RBF) kernels. In this way, a shape is represented as a classifier function, which has several advantages: (a)~the representation is completely generic, e.g. it can be applied to sparse landmark points, curves lines or a combination of the two, and~(b) it fuses inconsistent landmarks into consistent and directly comparable decision functions. Furthermore, this representation is also robust against noise, missing data and outliers \cite{Nguyen2013}.

@@

We extend the SVS representation to support also the case where multiple curves with different labels are annotated. This is useful when annotating different structures of the same object, such as the left and right eyebrows on faces. Even though not absolutely necessary in our framework, it can provide further guidance on the estimation of dense shape correspondences for various object classes. In more detail, we create a multichannel SVS image $\bm{d}(\bm{x})=[d_1(\bm{x}) \cdots d_i(\bm{x}) \cdots d_{N_c}(\bm{x})]$, where $d_i(\bm{x})$ is the SVS image that corresponds to the curve annotation of the $i$-th structure and $N_c$ is the total number of structures (a single curve annotation is the special case where $N_c=1$). Note that we do not necessarily require that all structures are annotated in all the object images: in the case that a structure is not annotated, the corresponding channel of the SVS image simply takes a zero value for all pixels. The shape flow estimation can deal with such missing information thanks to the spatial regularization and the low-rank constraint that it adopts, c.f.~Step~\ref{sec:step3}.

@@

Instead of the \emph{motion} basis used in multiframe optical flow formulation of \cite{Garg:2013hu}, we build a \emph{correspondence} basis that introduces constraints on how points of different shapes are matched to each other. Every pixel of the reference SVS image is matched to its corresponding position at every training SVS image and in this way defines a \emph{correspondence vector}. This vector consists of the 2D locations of the specific point in all SVS images. To form this vector, the training images are arranged in an arbitrary order. Similarly to the order of the training samples when PCA is applied, this order does not affect the result of our method and any re-ordering would produce exactly the same results.

@@

Using the constructed correspondence basis, the shape flow can be approximated as:
\vspace{-5pt}
\begin{equation}\label{eq:LinTrajModel}
    \bm{u}_n(\bx) \approx
    \textstyle \, \sum_{i=1}^R \, \bm{q}_i(n)v_i(\bx) \,\, ,
\vspace{-5pt}
\end{equation}
where $v_i(\bx)$ is the weight that needs to be applied on the $i$-th correspondence basis element, in order to get the correspondence vector for the point location $\bm{x}$. In other words, the shape flow for every point $\bm{x}$ is described as a linear combination of basis elements that is controlled by the coefficients $v_i(\bm{x})$.
The values of the $i$-th coefficient for all the points $v_i(\bm{x})$ can be interpreted as an image defined on $\Omega$. Using the correspondence basis, the determination of the shape flow boils down to the determination of the set of coefficients $v_i(\bm{x})$. The above representation of shape flow, constrains the correspondence vectors to lie on a subspace and, therefore, acts as a low-rank prior that enforces coherency of the shape registration result over the whole training dataset of shapes.

To effectively build the correspondence basis, we first transform the original annotations to sparse point clouds. Then, we apply the NICP algorithm of \cite{Amber2007} between the point cloud of annotations in the reference shape and the one of every shape of the training set. NICP iteratively deforms the cloud of points of every shape to match the points of the reference shape. This yields an initial estimation of the correspondence vectors on the sparse locations of annotated landmarks on the reference shape. Finally, the correspondence basis is found by applying PCA on these correspondence vectors and keeping only the first $R$ principal components.

@@

\begin{figure}[b!]
    \centering
    \includegraphics[width=0.9\columnwidth]{resources/models}
    \caption{Dense shape models build for faces and ears. Dense shapes are presented as grid for better visualization.}
    \label{fig:dense_models}
\end{figure}

@@

Let $\bm{d}(\bx;n)$, $\bm{d}(\bx;0):\Omega\rightarrow \R^{N_c}$ be the $n$-th training SVS image and the reference SVS image respectively. Following \cite{Garg:2013hu}, we propose to estimate the shape flow over all training images by minimizing the following energy:
\vspace{-5pt}
\begin{align}
E_{sf} & =\alpha
\int_{\Omega}\sum_{n=1}^{N_t} \|\bm{d}(\bx+\bm{u}_n(x);n)-\bm{d}(x;0)\| \ud \bx \label{eq:costfunc}\\
    &+ \beta \int_{\Omega}\sum_{n=1}^{N_t}\|\bm{u}_n(\bx)-\sum_{i=1}^R\bm{q}_i(n)v_i(\bx)\|^2 \ud \bx \label{eq:lowrank}\\
    &+
\int_\Omega  \sum_{i=1}^R \,\, \left \|    \nabla v_i(\bx)    \right \|  \,\ud \bx \label{eq:TVterm}
\vspace{-5pt}
\end{align}
This energy consists of two sets of unknown shape flows that are relatively close to each other: (i) $\bm{u}_n(x)$ which tries to explain the data from the input SVS images, and (ii) the shape flow determined by the correspondence basis coefficients $v_i(\bx)$ that are spatially regularised and enforce a low-rank prior.
%We minimize this energy jointly with respect to $\bm{u}_n(\bx)$ and $v_i(\bx)$.
%The positive parameters $\alpha$ and $\beta$ weigh the balance between the terms of the energy.

The \textbf{first term} of the above energy \eqref{eq:costfunc} is a data attachment term
that uses the robust $\Lone$-norm.  It is based on the assumption that the values of the reference SVS image $\bm{d}_0(\bx)$ at every pixel $\bx$ are preserved at its corresponding locations on all training SVS images $\bm{d}_n(\bx)$. The use of an $\Lone$-norm improves the robustness of the method since it allows deviations from this assumption, which might occur in practice.
The \textbf{second term} of the energy \eqref{eq:lowrank} penalizes the difference between the two sets  of shape flows and acts as a coupling term between them.
The \textbf{third term} of the energy \eqref{eq:TVterm} corresponds to the spatial Total Variation regularization \cite{rudin92} of
the correspondence basis coefficients $v_i(\bx)$.
This term penalizes spatial oscillations of each coefficient caused by distortions of the SVS images but not strong discontinuities that are desirable in the borders of different object regions. In addition, this term allows to fill in information into regions where the shape information in the SVS images is missing, due to e.g.~regions with no annotations.

@@

%In every coarse-to-fine and warping iteration, we use an initialization that comes from the previous iteration and for the very beginning, we initialize with a Thin Plate Splines (TPS) \cite{Bookstein1989} interpolation of the initial correspondence vectors described in Sec.~\ref{sec:step2}. We approximate the data term \eqref{eq:costfunc} by linearizing the SVS images around the initialization. After that, the energy becomes convex and we optimize it using alternating optimization w.r.t.~$v_i(\bx)$ and $\bm{u}_n(\bx)$. The minimization w.r.t.~
%$v_i(\bx)$ is decoupled for every coefficient $i$ and corresponds to Rudin-Osher-Fatemi Total Variation denoising \cite{rudin92}, which we solve efficiently by applying the first order primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}. The minimization w.r.t.~
%$\bm{u}_n(x)$ is decoupled for every pixel $\bx$ and every shape index $i$. This minimization is also implemented by applying the efficient primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}.

@@

Given a test image $\mathbf{I}$, the fitting process involves the minimization of the following cost function:
\begin{equation}
    \arg\min_{\bm{p}, \bm{c}}\|\bm{I}(\mathcal{W}(\bm{s}(\bm{p}))) - \bm{t}(\bm{c})\|_2^2
	\label{eq:aam_cost}
\end{equation}
This optimization problem is typically solved using the inverse-compositional Gauss-Newton algorithm, for which different variations have been proposed \cite{Matthews2004, Papandreou2008, Amberg2009, Tzimiropoulos2013, Alabort2014}. Note that the existence of the sampling function $\mathcal{W}()$ instead of a non-linear warping function has the advantage that all existing efficient gradient descent algorithms become exact.

\paragraph{Outline Patch-Based Active Appearance Model (PAAM)} The object classes for which the interior appearance does not have specific structure are modeled using patch-based AAMs \cite{Tzimiropoulos2014} trained on a set of sparse landmarks. Especially for human body parts (arms, legs), we strongly believe that the points located to the outline of the object are more suitable compared to the internal ones that correspond to the skeleton joints, which are commonly used by current literature \cite{buehler2011upper,charles2013domain,pfister2015flowing,yang2013articulated}.

The main differences between the patch-based and dense AAMs are that (a)~the densified shape instances are sub-sampled to include only the outline points, and (b)~the texture representation involves the sampling a neighbourhood around each point instead of a single pixel. Specifically, in order to build the outline sparse shape model, we simply select the outline points on the SVS reference frame. Then, by taking advantage of the dense correspondences obtained by Step \ref{sec:step3}, the shape model is trained in a similar way as in the dense case. Moreover, similar to the dense case, the texture model is built by sampling the image values from the sparse shape locations, i.e. $\mathcal{W}(\bm{s}(\bm{p}))$. However, contrary to dense AAMs, we sample a patch that is centred around each landmark point. These patches are then vectorised and concatenated in a single texture vector. Note that the optimization process remains exactly the same.

### 2016-04-09 contents/introduction.tex
cam ready final draft

\item Annotating with regards to consistent landmarks is an extremely time consuming, tedious and labour intensive work \cite{sagonas_iccv_300w_2013}, which is usually performed by a trained person. Furthermore, for various object classes, it requires a highly skilled person in order to identify and annotate landmarks in a consistent manner. For example, the human ear has very complicated inner structures (helix, crus antihelicis, scapha, tragus, lobe etc.) which remarkably differ between different ears. Moreover, certain ear parts, such as fossa triangularies and crus helicis, do not appear in all ears and their visibility is highly sensitive to the head pose and illumination variation. Another such example is the human body, which is generally annotated with regards to a number of landmarks that intuitively correspond to a set of body joints. For most body pose databases, the annotation task was undertaken by a crowd-sourcing Internet marketplace, so-called Amazon Mechanical Turk (AMT). Unfortunately, this resulted in acquiring inconsistent and inaccurate annotations, in many cases\footnote{In the case of faces, the quality of annotations produced from AMT are extremely inaccurate and cannot, by any means, be compared with the ones provided by the recent 300W competition \cite{sagonas_iccv_300w_2013,sagonas2016faces}.} (please see Figure \ref{fig:wrong_anno}). As it was also recently pointed out \cite{tompson2015efficient}, the inconsistencies in body joint annotations may also render the comparison between different human pose estimation methodologies irrelevant.

\item The nature of many deformable objects does not allow them to be annotated with regards to a consistent set of landmarks (e.g., bottles, fruits etc.). Additionally, it is very difficult to consistently annotate the boundary/outline of certain objects such as faces, ears, body, since these landmarks do not have any semantic meaning. That is why many state-of-the-art methods opt to leave the boundary out when reporting results \cite{Tzimiropoulos2014, Asthana2014}. The majority of the state-of-the-art methods for model-based landmark localisation \cite{Cao2012, Zhu2012, Xiong2013, Tzimiropoulos2014, Asthana2014} are not applicable to objects with inconsistent sets of landmarks.

@@

To illustrate how time consuming careful annotation of a complex deformable object is, we lay down our own experience based on the human ear. A trained annotator needs an average of 4 minutes per image for the manual annotation of 55 landmarks. This means that the annotation of 1000 images requires a total of about 67 hours. Furthermore, the quality of training as well as fatigue greatly influence the annotation accuracy. Hence, a second pass on the annotated data is, in many cases, necessary. Due to the fact that manual annotation is a costly and labour-intensive procedure, unsupervised learning of deformable models for the task of object alignment has recently attracted some attention \cite{frey2003learning, baker2004automatic, cootes2004groupwise, jojic2006escaping, Huang2006, kokkinos2007unsupervised, jiang2009learning, liu2009simultaneous, Zhang2012}. However, because the problem of fully unsupervised discovery of the deformations of arbitrary objects is difficult and ill-posed, the limited number of methods that have been proposed for the task cannot be directly applied to arbitrary collection of ``in-the-wild" images. On the other hand, the method of \cite{antonakos2014automatic}, which can deal with ``in-the-wild" images, requires a set of consistently annotated sparse shapes to perform deformable face congealing.

@@

Furthermore, we capitalise on recent advances on multiframe optical flow estimation \cite{Garg:2013hu,tomasi2012dense,snape15faceflow} and show that the relevant methodologies have matured enough to densely annotate the proposed shapes using either simplistic or even more sophisticated and robust shape representation methods \cite{Nguyen2013}.
In particular, in order to build dense correspondences between different shape instances of the same object class, we jointly estimate the optical flow among all the instances by imposing low-rank constrains, an approach that we call \emph{Shape Flow}. Multiframe optical flow has originally been applied on video sequences, relying on the assumptions of colour consistency and motion smoothness \cite{Garg:2013hu}. However, these assumptions do not hold in our case, where we have a collection of shapes. Therefore, we introduce appropriate modifications based on the consistency of image-based shape representation, as well as low rank priors.

@@

  \item We show that the employed patch-based SDM corrects the annotations that are currently provided for most major human body pose databases. Upon acceptance of the paper, we will make the corrected annotations publicly available.

>>>>

\item Annotating with regards to consistent landmarks is an extremely time-consuming, tedious and labour intensive work \cite{sagonas_iccv_300w_2013}, which is usually performed by a trained person. Furthermore, for various object classes, it requires a highly skilled person in order to identify and annotate landmarks in a consistent manner. For example, the human ear has very complicated inner structures (helix, crus antihelicis, scapha, tragus, lobe etc.) which remarkably differ between different ears. Moreover, certain ear parts, such as fossa triangularies and crus helicis, do not appear in all ears and their visibility is highly sensitive to the head pose and illumination variation. Another such example is the human body, which is generally annotated with regards to a number of landmarks that intuitively correspond to a set of body joints. For most body pose databases, the annotation task was undertaken by a crowd-sourcing Internet marketplace, so-called Amazon Mechanical Turk (AMT). Unfortunately, this resulted in acquiring inconsistent and inaccurate annotations, in many cases\footnote{In the case of faces, the quality of annotations produced from AMT are extremely inaccurate and cannot, by any means, be compared with the ones provided by the recent 300W competition \cite{sagonas_iccv_300w_2013,sagonas2016faces}.} (please see Figure \ref{fig:wrong_anno}). As it was also recently pointed out \cite{tompson2015efficient}, the inconsistencies in body joint annotations may also render the comparison between different human pose estimation methodologies irrelevant.

\item The nature of many deformable objects does not allow them to be annotated with regards to a consistent set of landmarks (e.g., bottles, fruits etc.). Additionally, it is very difficult to consistently annotate the outline of certain objects such as faces, ears, body, since these landmarks do not have any semantic meaning. That is why many state-of-the-art methods opt to leave the boundary out when reporting results \cite{Tzimiropoulos2014, Asthana2014}. The majority of the state-of-the-art methods for model-based landmark localisation \cite{Cao2012, Zhu2012, Xiong2013, Tzimiropoulos2014, Asthana2014} are not applicable to objects with inconsistent sets of landmarks.

@@

To illustrate how time-consuming careful annotation of a complex deformable object is, we lay down our own experience based on the human ear. A trained annotator needs an average of 4 minutes per image for the manual annotation of 55 landmarks. This means that the annotation of 1000 images requires a total of about 67 hours. Furthermore, the quality of training as well as fatigue greatly influence the annotation accuracy. Hence, a second pass on the annotated data is, in many cases, necessary. Due to the fact that manual annotation is a costly and labour-intensive procedure, unsupervised learning of deformable models for the task of object alignment has recently attracted some attention \cite{frey2003learning, baker2004automatic, cootes2004groupwise, jojic2006escaping, Huang2006, kokkinos2007unsupervised, jiang2009learning, liu2009simultaneous, Zhang2012}. However, because the problem of fully unsupervised discovery of the deformations of arbitrary objects is difficult and ill-posed, the limited number of methods that have been proposed for the task cannot be directly applied to arbitrary collections of ``in-the-wild" images. On the other hand, the method of \cite{antonakos2014automatic}, which can deal with ``in-the-wild" images, requires a set of consistently annotated sparse shapes to perform deformable face congealing.

@@

Furthermore, we capitalise on recent advances on multiframe optical flow estimation \cite{Garg:2013hu,tomasi2012dense,snape15faceflow} and show that the relevant methodologies have matured enough to densely annotate the proposed shapes using either simplistic or even more sophisticated and robust shape representation methods \cite{Nguyen2013}.
In particular, in order to build dense correspondences between different shape instances of the same object class, we jointly estimate the optical flow among all the instances by imposing low-rank constrains, an approach that we call \emph{Shape Flow}. Multiframe optical flow has originally been applied on video sequences, relying on the assumptions of colour consistency and motion smoothness \cite{Garg:2013hu}. However, these assumptions do not hold in our case, where we have a collection of shapes. Therefore, we introduce appropriate modifications based on the consistency of image-based shape representation, as well as low-rank priors.

@@

  \item We show that the employed patch-based SDM corrects the annotations that are currently provided for most major human body pose databases \footnote{\label{foot:annotations}The corrected annotations are publicly available in \url{http://www.ibug.doc.ic.ac.uk/resources/bodypose-anno-correction}.}.
